---
title: "Data Analysis"
output: html_document
date: "2023-11-13"
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}

# Set your working directory
library(here)
here()

# Loading packages
library(tidyverse)
library(lubridate)
library(cowplot)
library(sf)
library(leaflet)
library(mapview); mapviewOptions(fgb = FALSE)


# Reading in the data
SurgeUptakeNitrogen <-
  read.csv(here('Data','Processed','Nitrogen_Processed.csv'), stringsAsFactors=TRUE)

NitrogenYears <-read.csv(here('Data','Processed','NitrogenYears_Processed.csv'), stringsAsFactors=TRUE)
```

# Question 1: Does giant kelp nitrogen percent vary over year or season?

H0 : There is no significant difference of nitrogen percent over the past twenty years.
HA : There is significant difference of nitrogen percent over the past twenty years.

Prior to conducting this analysis, it was identified that nitrogen percent did not seem to visually change over the years. To get a better understanding..... 
```{r}
#1. LM for year
NitrogenRegressionYear <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Date)
#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .75 on 1280 degrees of freedom. However, year only explains 5% of nitrogen percent amount. 
summary(NitrogenRegressionYear)
#Correlation
cor.test(NitrogenYears$Nitrogen_Percent, NitrogenYears$Year)
#Running correlation test, the correlation score is -.202. So there is a negative correlation. So a somewhat weak correlation


#2. LM for Date
NitrogenRegressionDate <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Date)

#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .1648 on 728 degrees of freedom. Date explains 97.4% of nitrogen percent amount. 
summary(NitrogenRegressionDate)

#can't run corr because it is a date
#cor.test(NitrogenYears$Nitrogen_Percent, NitrogenYears$Date)

#3. Normal Distribution
par(mfrow = c(2,2), mar=c(2,2,2,2))
plot(NitrogenRegressionDate)
par(mfrow = c(1,1))

#Residuals v fitted, looking for symmetry. There is symmetry, and the line is fairly flat. There a re feww observations that are far off from the red line. Lowewr and higher ends do deviate from the Q-Q Residuals. Scale-location, there is very little symmetry with respects to the red line. Residuals vs Leverage, values outside dashed line show there are outliers. 
```


# Question 2: Does giant kelp nitrogen surge uptake vary per spring and summer season?
#Need to explain uptake rate calculation


```{r}
NitrogenSeasonBox <- 
  ggplot(Season, aes(x = Season, y = Uptake_Rate)) +
  geom_boxplot()+
  labs(title="Nitrogen Uptake Rate by Season", x="Season", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenSeasonBox)
ggsave(file=here('Output', 'Analysis','NitrogenSeasonBox.png'), width = 10, height = 10, units = "cm")

#BiomassSeasonBox <- 
#  ggplot(Season, aes(x = Biomass_Uptake, y = Season)) +
#  geom_boxplot()+
#  xlab("Biomass Uptake Rate")

#print(BiomassSeasonBox)


NitrogenSeasonPoint <- 
  ggplot(Season, aes(x = DayReplicate, y = Uptake_Rate, color=Season)) +
  geom_point(alpha=.6)+
  labs(title="Nitrogen Uptake Rate by Day and Season", x="Days", y="Nitrogen Uptake Rate (V(hr1))")


print(NitrogenSeasonPoint)
ggsave(file=here('Output', 'Analysis','NitrogenSeasonPoint.png'), width = 10, height = 10, units = "cm")

#BiomassSeasonPoint <- 
#  ggplot(Season, aes(x = DayReplicate, y = Biomass_Uptake, color=Season)) +
#  geom_point()+
#  xlab("Days of Study")+
#  ylab("Biomass Uptake Rate")

#print(BiomassSeasonPoint)

NitrogenSeasonBox2 <- 
  ggplot(Season, aes(x = Nitrogen_Types, y = Uptake_Rate, color=Season)) +
 geom_boxplot()+
 labs(title="Nitrogen Uptake Rate by Types and Season", x="Nitrogen Types", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenSeasonBox2)
ggsave(file=here('Output', 'Analysis','NitrogenSeasonBox2.png'), width = 10, height = 10, units = "cm")
```
```{r Season LM}
SeasonLM <-lm(data = Season, Uptake_Rate~Season)
summary(SeasonLM)

SeasonAOV <- aov(data=Season, Uptake_Rate~Season)
summary(SeasonAOV)


NitrogenAIC <- lm(data=Season, Uptake_Rate~Season+Nitrogen_Types+NewTime)
step(NitrogenAIC)

#really for uptake duration, not season
DurationAOV <- aov(data=Season, Uptake_Rate~NewTime)
summary(DurationAOV)

TukeyHSD(DurationAOV)

#NitrogenTukey<- HSD.test(DurationAOV, "Time", group=TRUE)
#print(NitrogenTukey)
```

# Question 3: Does giant kelp nitrogen surge uptake vary based on the amount of exposure?

```{r Q2 Exposure Length Plots}
ExposureLength <- SurgeUptakeNitrogen %>%
  mutate(Day_Replicate = as.factor(Day_Replicate))%>%
  mutate(Time = as.factor(Time))


ExposureLength$Season <-factor(Season$Season, labels = c("Spring", "Summer"))
ExposureLength$Nitrogen_Types<-factor(Season$Nitrogen_Types, labels= c("Ammonium", "Nitrate", "Urea"))


NitrogenTimeBox <- 
  ggplot(ExposureLength, aes(x = Time, y = Uptake_Rate)) +
  geom_boxplot()+
  labs(title="Nitrogen Uptake Rate by Exposure Length", x="Exposure Length (mins)", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenTimeBox)
ggsave(file=here('Output', 'Analysis','NitrogenTimeBox.png'), width = 10, height = 10, units = "cm")

#BiomassTimeBox <- 
#  ggplot(ExposureLength, aes(x = Time, y = Biomass_Uptake) +
#  geom_boxplot()+
#  xlab("Biomass Uptake Rate")

#print(BiomassTimeBox)


NitrogenTimePointbyDay <- 
  ggplot(ExposureLength, aes(x = Day_Replicate, y = Uptake_Rate, color=Time)) +
  geom_point(alpha=.6)+
  labs(title="Nitrogen Uptake Rate by Day and Exposure Length", x="Days", y="Nitrogen Uptake Rate (V(hr1))")


print(NitrogenTimePointbyDay)
ggsave(file=here('Output', 'Analysis','NitrogenTimePointbyDay.png'), width = 10, height = 10, units = "cm")

#BiomassTimePoint <- 
#  ggplot(ExposureLength, aes(x = Day_Replicate, y = Biomass_Uptake, color=Time)) +
#  geom_point()+
#  xlab("Days of Study")+
#  ylab("Biomass Uptake Rate")

#print(BiomassTimePoint)

NitrogenTimePointbyType <- 
  ggplot(ExposureLength, aes(x = Nitrogen_Types, y = Uptake_Rate, color=Time)) +
 geom_point()+
 labs(title="Nitrogen Uptake Rate by Types and Exposure Length", x="Nitrogen Types", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenTimePointbyType)
ggsave(file=here('Output', 'Analysis','NitrogenTimePointbyType.png'), width = 10, height = 10, units = "cm")
```
```{r Q2 Time LM}
TimeLM <-lm(data = ExposureLength, Uptake_Rate~Time)
summary(TimeLM)

TimeAOV <- aov(data=ExposureLength, Uptake_Rate~Time)
summary(TimeAOV)


TimeNitrogenAIC <- lm(data=ExposureLength, Uptake_Rate~Time+Nitrogen_Types+Season)
step(TimeNitrogenAIC)


TimeDurationAOV <- aov(data=ExposureLength, Uptake_Rate~Time)
summary(TimeDurationAOV)

TukeyHSD(TimeDurationAOV)

```

# Question 4: Does uptake vary by type of nitrogen (ammonium, nitrate, and urea)?

```{r}

# Creating a ggplot to compare uptake of various types of nitrogen

NitrogenUptaketype <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Nitrogen_Types, y = Uptake_Rate)) +
  geom_boxplot() +
  labs(title="Uptake Rate by Nitrogen Type", x="Nitrogen Types", y="Nitrogen Uptake Rate (V(hr1))")
print(NitrogenUptaketype)
ggsave(file=here('Output', 'Analysis','NitrogenUptaketype.png'), width = 10, height = 10, units = "cm")
# Creating a ggplot to compare biomass uptake of various types of nitrogen

NitrogenBiomassUptake <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Nitrogen_Types, y = Biomass_Uptake)) +
  geom_boxplot() +
  labs(title="Biomass Uptake Rate by Nitrogen Type", x="Nitrogen Types", y="Biomass Uptake Rate (µg/g/min)")
print(NitrogenBiomassUptake)
ggsave(file=here('Output', 'Analysis','NitrogenBiomassUptake.png'), width = 10, height = 10, units = "cm")

# Creating a ggplot to compare how nitrogen uptake varies per day per nitrogen type
NitrogenDayUptake <-
  ggplot(Season, aes(x = Nitrogen_Types, y = Uptake_Rate, color = DayReplicate)) +
  geom_jitter(alpha = 0.5) +
  scale_color_manual(values = c("#FF0000", "#BF003F", "#7F007F", "#3F00BF", "#0000FF"))+
  labs(title="Nitrogen Uptake per Day for Each Nitrogen Type", 
       x="Nitrogen Types", 
       y="Nitrogen Uptake Rate (V(hr1))", color = "Day")

# or
# NitrogenDayUptake <-
  #ggplot(SurgeUptakeNitrogen, aes(x = Day_Replicate, y = Uptake_Rate, color = Nitrogen_Types)) +
  #geom_point(alpha = 0.5) +
  #labs(title="Nitrogen Uptake per day for each type", 
      #x="Day", 
      # y="Nitrogen Uptake Rate (V(hr1))", color = "Nitrogen Types")

print(NitrogenDayUptake)
ggsave(file=here('Output', 'Analysis','NitrogenDayUptake.png'), width = 10, height = 10, units = "cm")

```

```{r Q3 LM}
# Generalized Linear Model (GLM)
NitorgenLM <-lm(data = SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types)
summary(NitorgenLM)

# One-way ANOVA
NitrogenAOV <- aov(data = SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types)
summary(NitrogenAOV)
TukeyHSD(NitrogenAOV)

# AIC-based Model Selection:

NitrogenTypeAIC <- lm(data=SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types+Season+Time)
step(NitrogenTypeAIC)
summary(NitrogenTypeAIC)

# For biomass
BiomassTypeAIC <- lm(data=SurgeUptakeNitrogen, Biomass_Uptake~Nitrogen_Types+Season+Time)
step(NitrogenTypeAIC)
summary(NitrogenTypeAIC)

```

