---
title: "Data Analysis"
output: html_document
date: "2023-11-13"
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE}

# Set your working directory
library(here)
here()

#install.packages("forcats")

# Loading packages
library(tidyverse)
library(lubridate)
library(cowplot)
library(agricolae)
library(sf)
library(leaflet)
library(mapview); mapviewOptions(fgb = FALSE)
library(forcats)


# Reading in the data
SurgeUptakeNitrogen <-
  read.csv(here('Data','Processed','Nitrogen_Processed.csv'), stringsAsFactors=TRUE)

NitrogenYears <-read.csv(here('Data','Processed','NitrogenYears_Processed.csv'), stringsAsFactors=TRUE)
```

# Question 1:	Has nitrogen concentration in giant kelp changed over the years (2002-2021)?

H0 : There is no significant difference of nitrogen percent over the past twenty years.
HA : There is significant difference of nitrogen percent over the past twenty years.

Prior to conducting this analysis, it was identified that nitrogen percent did not seem to visually change over the years. To get a better understanding of the relationship, linear regression and time series analysis was conducted.  
```{r Q1}
#1. LM for year
NitrogenRegressionYear <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Year)
#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .75 on 1280 degrees of freedom. However, year only explains 5% of nitrogen percent amount. 
summary(NitrogenRegressionYear)
#Correlation
cor.test(NitrogenYears$Nitrogen_Percent, NitrogenYears$Year)
#Running correlation test, the correlation score is -.202. So there is a negative correlation. So a somewhat weak correlation

par(mfrow = c(2,2), mar=c(2,2,2,2))
plot(NitrogenRegressionYear)
par(mfrow = c(1,1))


#2. LM for Date
NitrogenRegressionDate <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Date)

#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .1648 on 728 degrees of freedom. Date explains 97.4% of nitrogen percent amount. 
summary(NitrogenRegressionDate)

#can't run corr because it is a date
#cor.test(NitrogenYears$Nitrogen_Percent, NitrogenYears$Date)

#3. Normal Distribution
par(mfrow = c(2,2), mar=c(2,2,2,2))
plot(NitrogenRegressionDate)
par(mfrow = c(1,1))

#Residuals v fitted, looking for symmetry. There is symmetry, and the line is fairly flat. There a re few observations that are far off from the red line. Lower and higher ends do deviate from the Q-Q Residuals. Scale-location, there is very little symmetry with respects to the red line. Residuals vs Leverage, values outside dashed line show there are outliers. 


#4. Scatter plot
#nitrogenbydate <- ggplot(NitrogenYears, aes(x=Date, y=Nitrogen_Percent))+
 # geom_point()+
 # geom_smooth(method="lm")

#why is geomsmooth not showing up?
#how to get less dates on the x axis?
#print(nitrogenbydate)

#5. Tukey
#NitrogenDateAOV <- aov(data = NitrogenYears, Nitrogen_Percent ~Date)
#summary(NitrogenDateAOV)

#TukeyHSD(NitrogenDateAOV)

#NitrogenDateTotals <- HSD.test(NitrogenDateAOV, "Year", group =TRUE)
#NitrogenDateTotals

NitrogenYears$Year <- as.factor(NitrogenYears$Year)
NitrogenPlot <- ggplot(NitrogenYears, aes(x=Year, y=Nitrogen_Percent))+
  geom_boxplot()
print(NitrogenPlot)



```
```{r}
#6. Season Plot #average the months
Nitrogen.monthly.preliminary <- NitrogenYears %>%
  mutate(CleanDate = my(paste0(Month,"-",Year)))

Nitrogen.monthly <-aggregate(Nitrogen.monthly.preliminary$Nitrogen_Percent,  
                                  by=list(Nitrogen.monthly.preliminary$CleanDate),  
                                  FUN=mean)

colnames(Nitrogen.monthly) <- c("Date", "Mean_NitrogenPercent")

#7. Generate time series monthly
f_monthmonthly <- month(first(Nitrogen.monthly$Date))
f_yearmonthly <- year(first(Nitrogen.monthly$Date))

Nitrogen.monthly.ts <- ts(Nitrogen.monthly$Mean_NitrogenPercent,
                   start=c(f_yearmonthly,f_monthmonthly),
                   frequency=12) 

#Decompose monthly
Nitrogen.monthly.ts.decomp <- stl(Nitrogen.monthly.ts,s.window = "periodic")
plot(Nitrogen.monthly.ts.decomp)

#12 Run SMK test monthly
Nitrogen_Monthly_SMK <- Kendall::SeasonalMannKendall(Nitrogen.monthly.ts)

Nitrogen_Monthly_SMK
summary(Nitrogen_Monthly_SMK)

# 13 Visualization
Nitrogen_Plot <-
ggplot(Nitrogen.monthly, aes(x = Date, y = Mean_NitrogenPercent)) +
  geom_point() +
  geom_line() +
  ylab("Mean Nitrogen (percent)") +
  geom_smooth( method = lm )

print(Nitrogen_Plot)

#taking out seasonal
NitrogenComponents <- as.data.frame(Nitrogen.monthly.ts.decomp$time.series[,1:3])

Nitrogen_NoSeasonal <-NitrogenComponents$trend+NitrogenComponents$remainder

#16 Run MK test monthly
Nitrogen_Monthly_MK <- Kendall::MannKendall(Nitrogen_NoSeasonal)

Nitrogen_Monthly_MK
summary(Nitrogen_Monthly_MK)
```
# Question 2:	Does giant kelp nitrogen concentration amount vary by season?
H0 : There is no significant difference of nitrogen concentration by season.
HA : There is significant difference of nitrogen concentration by season.
```{r Q1 cont.}
#More month stuff

NitrogenYears$Month <- factor(NitrogenYears$Month, levels = c(1,2,3,4,5,6,7,8,9,10,11,12),labels =c(  
    "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep",  
    "Oct", "Nov", "Dec") )

NitrogenYears <- NitrogenYears %>%
  mutate(SeasonNames = fct_collapse(Month,
    "Spring" = month.abb[3:5],
    "Summer" = month.abb[6:8],
    "Fall" = month.abb[9:11],
    "Winter" = month.abb[c(12,1,2)]))


NitrogenRegressionSeason <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$SeasonNames)
#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .75 on 1280 degrees of freedom. However, year only explains 5% of nitrogen percent amount. 
summary(NitrogenRegressionSeason)

NitrogenRegressionMonth <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Month)
summary(NitrogenRegressionMonth)


#3. Normal Distribution
par(mfrow = c(2,2), mar=c(2,2,2,2))
plot(NitrogenRegressionSeason)
par(mfrow = c(1,1))

#Residuals v fitted, looking for symmetry. There is symmetry, and the line is fairly flat. There a re feww observations that are far off from the red line. Lowewr and higher ends do deviate from the Q-Q Residuals. Scale-location, there is very little symmetry with respects to the red line. Residuals vs Leverage, values outside dashed line show there are outliers. 


NitrogenSeasonPlot <- ggplot(NitrogenYears, aes(x=SeasonNames, y=Nitrogen_Percent))+
  geom_boxplot()
print(NitrogenSeasonPlot)


```


# Question 3:	Does nitrogen surge uptake vary between the spring and summer seasons?
#Need to explain uptake rate calculation
H0 : There is no significant difference of nitrogen surge uptake between spring and summer seasons.
HA : There is significant difference of nitrogen surge uptake between spring and summer seasons.

The metadata indicates that the giant kelp blades studied were collected during the nitrate replete spring and the nitrate-deplete summer. The five replicate experiments were conducted in both spring and summer of 2019, which allows us to find potential differences in the nitrogen surge uptake of the two populations. The experimenters calculate uptake rate using a modified version of an equation by Legendre and Grosselin (1997) V= (nt - n0)/[t(ds-d0)] where nt is the atom percentage of Nitrogen-15 in the blade after incubation; n0 is the average concentration of isotope in the control blades; ds is the atom percentage of Nitrogen-15 in the seawater at the start of the incubation; d0 is the atom percentage of Nitrogen-15 in the nitrate pool before isotope tracer was added; and t is the length of the incubation time in minutes. 

```{r}
NitrogenSeasonBox <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Season, y = Uptake_Rate)) +
  geom_boxplot()+
  labs(title="Nitrogen Uptake Rate by Season", x="Season", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenSeasonBox)
#ggsave(file=here('Output', 'Analysis','NitrogenSeasonBox.png'), width = 10, height = 10, units = #"cm")

#BiomassSeasonBox <- 
#  ggplot(Season, aes(x = Biomass_Uptake, y = Season)) +
#  geom_boxplot()+
#  xlab("Biomass Uptake Rate")

#print(BiomassSeasonBox)


NitrogenSeasonPoint <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Day_Replicate, y = Uptake_Rate, color=Season)) +
  geom_point(alpha=.5)+
  labs(title="Nitrogen Uptake Rate by Day and Season", x="Days", y="Nitrogen Uptake Rate (V(hr1))")


print(NitrogenSeasonPoint)
#ggsave(file=here('Output', 'Analysis','NitrogenSeasonPoint.png'), width = 10, height = 10, units = #"cm")

#BiomassSeasonPoint <- 
#  ggplot(Season, aes(x = DayReplicate, y = Biomass_Uptake, color=Season)) +
#  geom_point()+
#  xlab("Days of Study")+
#  ylab("Biomass Uptake Rate")

#print(BiomassSeasonPoint)

NitrogenSeasonBox2 <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Nitrogen_Types, y = Uptake_Rate, color=Season)) +
 geom_boxplot()+
 labs(title="Nitrogen Uptake Rate by Types and Season", x="Nitrogen Types", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenSeasonBox2)
#ggsave(file=here('Output', 'Analysis','NitrogenSeasonBox2.png'), width = 10, height = 10, units = #"cm")



#linear regression
NitrogenRegressionSeason <- lm(SurgeUptakeNitrogen$Uptake_Rate ~ SurgeUptakeNitrogen$Season)

summary(NitrogenRegressionSeason)


```
```{r Season LM}
SurgeUptakeNitrogen$Time <- as.factor(SurgeUptakeNitrogen$Time)

SeasonLM <-lm(data = SurgeUptakeNitrogen, Uptake_Rate~Season)
summary(SeasonLM)

SeasonAOV <- aov(data=SurgeUptakeNitrogen, Uptake_Rate~Season)
summary(SeasonAOV)


NitrogenAIC <- lm(data=SurgeUptakeNitrogen, Uptake_Rate~Season+Nitrogen_Types+Time)
step(NitrogenAIC)

#really for uptake duration, not season
#DurationAOV <- aov(data=SurgeUptakeNitrogen, Uptake_Rate~Time)
#summary(DurationAOV)

#TukeyHSD(DurationAOV)
#TukeyHSD(SeasonAOV)

#NitrogenTukey<- HSD.test(DurationAOV, "Time", group=TRUE)
#print(NitrogenTukey)
```

# Question 4: Does giant kelp nitrogen surge uptake vary based on the amount of exposure?
H0 : There is no significant difference of nitrogen surge uptake between 
different exposure lengths.
HA : There is significant difference of nitrogen surge uptake between different
exposure lengths.
```{r Q3 Exposure Length Plots}
ExposureLength <- SurgeUptakeNitrogen %>%
  mutate(Day_Replicate = as.factor(Day_Replicate))%>%
  mutate(Time = as.factor(Time))


ExposureLength$Season <-factor(SurgeUptakeNitrogen$Season, labels = c("Spring", "Summer"))
ExposureLength$Nitrogen_Types<-factor(SurgeUptakeNitrogen$Nitrogen_Types, labels= c("Ammonium", "Nitrate", "Urea"))


NitrogenTimeBox <- 
  ggplot(ExposureLength, aes(x = Time, y = Uptake_Rate)) +
  geom_boxplot()+
  labs(title="Nitrogen Uptake Rate by Exposure Length", x="Exposure Length (mins)", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenTimeBox)
ggsave(file=here('Output', 'Analysis','NitrogenTimeBox.png'), width = 10, height = 10, units = "cm")

#BiomassTimeBox <- 
#  ggplot(ExposureLength, aes(x = Time, y = Biomass_Uptake) +
#  geom_boxplot()+
#  xlab("Biomass Uptake Rate")

#print(BiomassTimeBox)


NitrogenTimePointbyDay <- 
  ggplot(ExposureLength, aes(x = Day_Replicate, y = Uptake_Rate, shape=Time)) +
  geom_point(alpha=.2)+
  labs(title="Nitrogen Uptake Rate by Day and Exposure Length", x="Days", y="Nitrogen Uptake Rate (V(hr1))")


print(NitrogenTimePointbyDay)
ggsave(file=here('Output', 'Analysis','NitrogenTimePointbyDay.png'), width = 10, height = 10, units = "cm")

#BiomassTimePoint <- 
#  ggplot(ExposureLength, aes(x = Day_Replicate, y = Biomass_Uptake, color=Time)) +
#  geom_point()+
#  xlab("Days of Study")+
#  ylab("Biomass Uptake Rate")

#print(BiomassTimePoint)

NitrogenTimePointbyType <- 
  ggplot(ExposureLength, aes(x = Nitrogen_Types, y = Uptake_Rate, shape=Time)) +
 geom_point(alpha=.2)+
 labs(title="Nitrogen Uptake Rate by Types and Exposure Length", x="Nitrogen Types", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenTimePointbyType)
ggsave(file=here('Output', 'Analysis','NitrogenTimePointbyType.png'), width = 10, height = 10, units = "cm")
```
```{r Q3 Time LM}
TimeLM <-lm(data = ExposureLength, Uptake_Rate~Time)
summary(TimeLM)

TimeAOV <- aov(data=ExposureLength, Uptake_Rate~Time)
summary(TimeAOV)


TimeNitrogenAIC <- lm(data=ExposureLength, Uptake_Rate~Time+Nitrogen_Types+Season)
step(TimeNitrogenAIC)


TimeDurationAOV <- aov(data=ExposureLength, Uptake_Rate~Time)
summary(TimeDurationAOV)

TukeyHSD(TimeDurationAOV)

```

# Question 5: Does uptake vary by type of nitrogen (ammonium, nitrate, and urea)?
H0 : There is no significant difference of nitrogen surge uptake of different nitrogen types.
HA : There is significant difference of nitrogen surge uptake between different nitrogen types.

```{r Q4}

# Creating a ggplot to compare uptake of various types of nitrogen

NitrogenUptaketype <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Nitrogen_Types, y = Uptake_Rate)) +
  geom_boxplot() +
  labs(title="Uptake Rate by Nitrogen Type", x="Nitrogen Types", y="Nitrogen Uptake Rate (V(hr1))")
print(NitrogenUptaketype)
ggsave(file=here('Output', 'Analysis','NitrogenUptaketype.png'), width = 10, height = 10, units = "cm")
# Creating a ggplot to compare biomass uptake of various types of nitrogen

NitrogenBiomassUptake <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Nitrogen_Types, y = Biomass_Uptake)) +
  geom_boxplot() +
  labs(title="Biomass Uptake Rate by Nitrogen Type", x="Nitrogen Types", y="Biomass Uptake Rate (µg/g/min)")
print(NitrogenBiomassUptake)
ggsave(file=here('Output', 'Analysis','NitrogenBiomassUptake.png'), width = 10, height = 10, units = "cm")

# Creating a ggplot to compare how nitrogen uptake varies per day per nitrogen type
NitrogenDayUptake <-
  ggplot(Season, aes(x = Nitrogen_Types, y = Uptake_Rate, color = DayReplicate)) +
  geom_jitter(alpha = 0.2) +
  scale_color_manual(values = c("#FF0000", "#BF003F", "#7F007F", "#3F00BF", "#0000FF"))+
  labs(title="Nitrogen Uptake per Day for Each Nitrogen Type", 
       x="Nitrogen Types", 
       y="Nitrogen Uptake Rate (V(hr1))", color = "Day")

# or
# NitrogenDayUptake <-
  #ggplot(SurgeUptakeNitrogen, aes(x = Day_Replicate, y = Uptake_Rate, color = Nitrogen_Types)) +
  #geom_point(alpha = 0.5) +
  #labs(title="Nitrogen Uptake per day for each type", 
      #x="Day", 
      # y="Nitrogen Uptake Rate (V(hr1))", color = "Nitrogen Types")

print(NitrogenDayUptake)
ggsave(file=here('Output', 'Analysis','NitrogenDayUptake.png'), width = 10, height = 10, units = "cm")

```

The box plot below shows the change in biomass uptake rate by nitrogen type.
```{r, echo=FALSE, message=FALSE, fig.cap="Biomass Uptake by Type Box Plot"}
NitrogenBiomassUptake <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Nitrogen_Types, y = Biomass_Uptake)) +
  geom_boxplot() +
  labs(title="Biomass Uptake Rate by Nitrogen Type", x="Nitrogen Types", y="Biomass Uptake Rate (µg/g/min)")
print(NitrogenBiomassUptake)
ggsave(file=here('Output', 'Analysis','NitrogenBiomassUptakeBoxNitrogenType.png'), width = 10, height = 10, units = "cm")
```
```

```{r Q4 LM}
# Generalized Linear Model (GLM)
NitorgenLM <-lm(data = SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types)
summary(NitorgenLM)

# One-way ANOVA
NitrogenAOV <- aov(data = SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types)
summary(NitrogenAOV)
TukeyHSD(NitrogenAOV)

# AIC-based Model Selection:

NitrogenTypeAIC <- lm(data=SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types+Season+Time)
step(NitrogenTypeAIC)
summary(NitrogenTypeAIC)

# For biomass
BiomassTypeAIC <- lm(data=SurgeUptakeNitrogen, Biomass_Uptake~Nitrogen_Types+Season+Time)
step(NitrogenTypeAIC)
summary(NitrogenTypeAIC)

```

