---
title: "Trends in Giant Kelp Nitrogen Concentration and Surge Uptake"
author: "Erin Ansbro, Cara Kuuskvere, Tani Valdez Rivas"
date: "December 2023"
output: pdf_document
---

```{r include=FALSE}
# Set your working directory
library(here)
here()

# Loading packages
library(tidyverse, quietly=TRUE)
library(lubridate)
library(cowplot)
library(knitr)
library(sf)
library(leaflet)
#library(mapview); mapviewOptions(fgb = FALSE)


# Reading in the data
SurgeUptakeNitrogen <-
  read.csv(here('Data','Processed','Nitrogen_Processed.csv'), stringsAsFactors=TRUE)

# Set your ggplot theme
#Create and set theme 
mytheme <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
theme_set(mytheme)

# Load your datasets
SurgeUptakeNitrogen <-
  read.csv(here('Data','Processed','Nitrogen_Processed.csv'), stringsAsFactors=TRUE)

NitrogenYears <-read.csv(here('Data','Processed','NitrogenYears_Processed.csv'), stringsAsFactors=TRUE)

#Verifying data attributes
SurgeUptakeNitrogen$Time <- as.factor(SurgeUptakeNitrogen$Time)
SurgeUptakeNitrogen$Blade_Replicate <- as.factor(SurgeUptakeNitrogen$Blade_Replicate)
SurgeUptakeNitrogen$Day_Replicate <- as.factor(SurgeUptakeNitrogen$Day_Replicate)

```
\newpage
\tableofcontents
\newpage
\listoftables 
\newpage
\listoffigures
\newpage


# Rationale and Research Questions

Kelp forests along the United States West Coast are impacted by increasing temperatures as a result of climate change. Since the late 19th century, Santa Barbara County in Southern California has had an increase in average temperature by 4.1 degrees Fahrenheit (Wilson, 2019). Kelp not only provides a healthy ecosystem for other flora and fauna but also provides ecosystem services to humans as well. Kelp assists with human-related environmental issues in two ways. First, kelp can sequester a large amount of carbon, aiding in climate change mitigation (Hurlimann, 2019). Secondly, they absorb excess nitrogen, which can lessen the impacts of agricultural runoff (Conroy, 2023).   

A recent study has shown that increased nitrogen uptake by kelp can limit the effect that heat has on them. This is an interesting development because nitrogen and eutrophication are generally seen as negative consequences of human interference. However, in this case, it could assist kelp and help them survive increased climate-change-related temperature increases (Fernández, 2020).
Interested in this recent research, data about nitrogen concentration amount and uptake were located in the Long-Term Ecological Research database titled “SBC LTER: Reef: Surge uptake capability in Macrocystis pyrifera in response to pulses of three different forms of nitrogen” (Cedeno, 2021) and ” SBC LTER: Reef: Macrocystis pyrifera CHN content (carbon, hydrogen, nitrogen), ongoing since 2002” (Santa Barbara, 2021). This data looked simple enough to clean and manipulate, while also providing enough information for a research project. 

We are interested in trends in giant kelp nitrogen concentration amount and nitrogen surge uptake. Specific questions include:
1.	Has nitrogen concentration in kelp changed over the years (2002-2021)?
2.	Does giant kelp nitrogen concentration amount vary by season?
3.	Does nitrogen surge uptake vary between the spring and summer seasons?
4.	Does giant kelp nitrogen surge uptake vary based on the amount of exposure?
5.	Does uptake vary by type of nitrogen (ammonium, nitrate, and urea)?

Github Information: https://github.com/Tani-ValdezRivas/ValdezKuuskvereAnsbro


\newpage

# Dataset Information

Data sets were found on the Environmental Data Initiative. All sets were from the Santa Barbara Coastal Long Term Ecological Research Site (SBC LTER) which is based  within the University of California, Santa Barbara (UCSB) Marine Science Institute, and is part of the National Science Foundation's LTER Network. 

The first data set focuses on one experiment done that examines nitrogen and biomass uptake by kelp during timed nitrogen blasting.  The first data set was used to determine if nitrogen uptake varies by time (minutes), season, or nitrogen type. The second data set has dates and percent nitrogen of kelp over almost a twenty year period. This second data set was used to gain a general sense of nitrogen found in kelp in Santa Barbara, CA and to see if that has changed over time.


Table:Data Set Information

Detail | Data Set 1 | Data Set 2
-----:|:----:|:----:
Source| Santa Barbara Coastal Long Term Ecological Research (LTER) | Santa Barbara Coastal  Long Term Ecological Research (LTER)
Retrieved from | https://portal.edirepository.org/nis/mapbrowse?packageid=edi.1022.2 |https://doi.org/10.6073/pasta/48686fef076a23310a0e48f69d76260c 
Variables Used | Nitrogen Types, Season, Day Replicate, Time, Blade Replicate, Uptake Rate, Biomass Uptake Rate |Year, Month, Date, Nitrogen Percent
Date | 2021 | 2002-2021

The data for nitrogen uptake was cleaned by renaming the column names to be better understood by the public. Multiple columns including columns for Blade Replicate, Day Replicate, Time, Season, and Nitrogen Type were changed to factors. Season and Nitrogen Type also had labels added to capitalize the names of seasons and nitrogen types.Data cleaning for the second data set entailed renaming column names as well. Date was changed from a factor to a date data type. Sites with missing values were also removed. 

\newpage

# Exploratory Analysis 

The first part of the exploratory analysis focused on the dataset from the research study conducted on giant kelp over five days. Box plots were used to explore the relationship of the variables (Nitrogen Type, Season, Length of Exposure, Day Replicate, and Blade Replicate) against nitrogen uptake rate. 


The metadata indicates that the giant kelp blades studied were collected during the nitrate replete spring and the nitrate-deplete summer. The five replicate experiments were conducted in both spring and summer of 2019, which allows us to find potential differences in the nitrogen surge uptake of the two populations. The experimenters calculate uptake rate using a modified version of an equation by Legendre and Grosselin (1997) V= (nt - n0)/[t(ds-d0)] where nt is the atom percentage of Nitrogen-15 in the blade after incubation; n0 is the average concentration of isotope in the control blades; ds is the atom percentage of Nitrogen-15 in the seawater at the start of the incubation; d0 is the atom percentage of Nitrogen-15 in the nitrate pool before isotope tracer was added; and t is the length of the incubation time in minutes.


```{r Box Plots, echo=FALSE, message=FALSE, fig.cap="Uptake Rate by Nitrogen Type Box Plot"}
# Data Exploration Part 2.1: Box Plots to visualize data means and quartiles
UptakeRateBoxbyType <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Nitrogen_Types, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Nitrogen Type", x="Nitrogen Types", y="Nitrogen Uptake Rate(V(hr1))")
UptakeRateBoxbyType
ggsave(file=here('Output', 'BoxDataExp','UptakeRateBoxbyType.png'), width = 10, height = 10, units = "cm")

```


```{r, echo=FALSE, message=FALSE, fig.cap="Uptake Rate by Season Box Plot"}
UptakeRateBoxbySeason <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Season, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Season", y="Nitrogen Uptake Rate (V(hr1))")
UptakeRateBoxbySeason
ggsave(file=here('Output', 'BoxDataExp','UptakeRateBoxbySeason.png'), width = 10, height = 10, units = "cm")
```

```{r, echo=FALSE, message=FALSE, fig.cap="Uptake Rate byExposure Length Box Plot"}
UptakeRateBoxbyTime <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Time, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Length of Exposure", x="Length of Exposure (minutes)", y="Nitrogen Uptake Rate (V(hr1))")
UptakeRateBoxbyTime
ggsave(file=here('Output','BoxDataExp', 'UptakeRateBoxbyTime.png'), width = 10, height = 10, units = "cm")

```


```{r, echo=FALSE, message=FALSE, fig.cap="Uptake Rate Day Replicate Box Plot"}

UptakeRateBoxbyDayRep <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Day_Replicate, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Day Replicate", x="Day Replicate", y="Nitrogen Uptake Rate Uptake Rate (V(hr1))")

UptakeRateBoxbyDayRep
ggsave(file=here('Output', 'BoxDataExp','UptakeRateBoxbyDayRep.png'), width = 10, height = 10, units = "cm")
```



```{r, echo=FALSE, message=FALSE, fig.cap="Nitrogen Uptake Rate by Blade Replicate Box Plot"}

UptakeRateBoxbyBladeRep <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Blade_Replicate, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Blade Replicate", x="Blade Replicate", y="Nitrogen Uptake Rate (V(hr1))")
UptakeRateBoxbyBladeRep
ggsave(file=here('Output','BoxDataExp', 'UptakeRateBoxbyBladeRep.png'), width = 10, height = 10, units = "cm")
```
Violin plots were used to understand the density of the variables when comparing with nitrogen uptake rate. See five violin plots below. 


```{r Violins, echo=FALSE, message=FALSE, fig.cap="Nitrogen Uptake Rate by Nitrogen Types Violin Plot"}
# Data Exploration Part 2.2: Violins to visualize range and spread of each component of the data set
UptakeRateViolinbyType <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Nitrogen_Types, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Nitrogen Type", x="Nitrogen Types", y="Nitrogen Uptake Rates (V(hr1)")
UptakeRateViolinbyType
ggsave(file=here('Output','ViolinDataExp', 'UptakeRateViolinbyType.png'), width = 10, height = 10, units = "cm")

```



```{r, echo=FALSE, message=FALSE, fig.cap="Nitrogen Uptake Rate by Season Violin Plot"}

UptakeRateViolinbySeason <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Season, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Season", y="Nitrogen Uptake Rates (V(hr1)", x="Season")
UptakeRateViolinbySeason
ggsave(file=here('Output','ViolinDataExp', 'UptakeRateViolinbySeason.png'), width = 10, height = 10, units = "cm")
```

```{r, echo=FALSE, message=FALSE, fig.cap="Nitrogen Uptake Rate by Length of Exposure Violin Plot"}

UptakeRateViolinbyTime <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Time, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Length of Exposure", y="Nitrogen Uptake Rate (V(hr1))", x="Length of Exposure (minutes)")
UptakeRateViolinbyTime
ggsave(file=here('Output', 'ViolinDataExp','UptakeRateViolinbyTime.png'), width = 10, height = 10, units = "cm")

```


```{r, echo=FALSE, message=FALSE, fig.cap="Nitrogen Uptake Rate by Day Replicate Violin Plot"}

UptakeRateViolinbyDayRep <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Day_Replicate, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Day Replicate", x="Day Replicate", y="Nitrogen Uptake Rate (V(hr1))")
UptakeRateViolinbyDayRep
ggsave(file=here('Output', 'ViolinDataExp','UptakeRateViolinbyDayRep.png'), width = 10, height = 10, units = "cm")
```

```{r, echo=FALSE, message=FALSE, fig.cap="Nitrogen Uptake Rate by Blade Replicate Violin Plot"}

UptakeRateViolinbyBladeRep <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Blade_Replicate, y = Uptake_Rate))+
  labs(title="Nitrogen Uptake Rate (V(hr1)) by Blade Replicate", x="Blade Replicate", y="Nitrogen Uptake Rate (V(hr1))")
UptakeRateViolinbyBladeRep
ggsave(file=here('Output', 'ViolinDataExp','UptakeRateViolinbyBladeRep.png'), width = 10, height = 10, units = "cm")
```


All of the day replicates and blade replicate nitrogen uptake rates were used when comparing the variables of Season, Length of Exposure, and Nitrogen Type were studied further. From both the box plot and violin plot, nitrogen uptake was high in the first minute. Summer also had slightly more and higher nitrogen uptake rates. Ammonium also had a higher range of nitrogen uptake rates than the other two nitrogen types. All three of these variables were studied further to understand if they contributed to a significant different of nitrogen uptake rate. 

\newpage
The second part of the exploratory analysis focused on an overview of nitrogen concentration amount over the last twenty years. A summary was used on the entire data set as well as on the nitrogen concentration column to gain a general sense of what amounts were present.  
```{r, echo=FALSE, message=FALSE}
#Data Exploration Part 3

summary(NitrogenYears)

summary(NitrogenYears$Nitrogen_Percent)
```
  
  A scatter plot was used to visualize the relationship between year and nitrogen concentration amount.
```{r,  message=FALSE, echo=FALSE, fig.cap="Nitrogen Concentration by Year Scatter Plot"}
#Data Exploration Part 3

NitrogenYears_Point <- ggplot(NitrogenYears, aes(x=Year, y=Nitrogen_Percent))+
  geom_point(alpha=.5)+
  labs(title="Nitrogen Percent by Year", x="Date", y="Nitrogen Percent")

ggsave(file=here('Output', 'ExploratoryAnalysis','NitrogenConcentrationBoxbySeason.png'), width = 10, height = 10, units = "cm")


print(NitrogenYears_Point)
```

  
  Lastly, a boxplot, see below, was used to visualize the relationship between sites and nitrogen concentration. Looking at this data, it appeared concentration amount was not too different by site. Looking at the scatter plot for year, it looked like nitrogen decreased slightly. This was then explored further in the Data Analysis section.

```{r, message=FALSE, fig.cap="Exploratory Analysis for Nitrogen Concentration by Site"}
#Data Exploration Part 3

NitrogenYears_Box <- ggplot(NitrogenYears, aes(x=NewSites, y=Nitrogen_Percent))+
  geom_boxplot()+
  labs(title="Nitrogen Percent by Site", x="Site", y="Nitrogen Percent")

ggsave(file=here('Output', 'ExploratoryAnalysis','NitrogenConcentrationBoxbySite.png'), width = 10, height = 10, units = "cm")

print(NitrogenYears_Box)
```

\newpage

# Analysis


## Question 1: Does giant kelp nitrogen percent vary over year or season?

For the first question, we wanted to understand the relationship between nitrogen concentration and time. Linear regression and normal distribution tests were run for both the year and the specific date of collection. The box plot below shows the nitrogen concentrations over the years, starting with 2002. 

H0 : There is no significant difference of nitrogen percent over the past twenty years.
HA : There is significant difference of nitrogen percent over the past twenty years.

The interaction between nitrogen concentration and year is significant (p < 2.2e-16). The interaction between nitrogen concentration and collection date is significant (p < 0.001). We can reject the null hypothesis and accept that there is significant difference of nitrogen percent from 2002 to 2021. The R-squared is .974, so a high amount of nitrogen concentration variance is in response to date change. Below is a box plot showing change in nitrogen concentration over time.


```{r, message=FALSE, fig.cap="Nitrogen Concentration over the Years"}
#2. LM for Date
NitrogenRegressionDate <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Date)

#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .1648 on 728 degrees of freedom. Date explains 97.4% of nitrogen percent amount. 
summary(NitrogenRegressionDate)


NitrogenYears$Year <- as.factor(NitrogenYears$Year)

NitrogenPlot <- ggplot(NitrogenYears, aes(x=Year, y=Nitrogen_Percent))+
  geom_boxplot()+
  labs(title="Nitrogen Concentration Percent by Year", x="Year", y="Nitrogen Concentration Percent")
print(NitrogenPlot)

ggsave(file=here('Output', 'Analysis','NitrogenConcentrationBoxbyDate.png'), width = 10, height = 10, units = "cm")

```

  
  Below are the normal distribution residuals charts for the linear regression of nitrogen concentration by date.
```{r, message=FALSE, fig.cap="Nitrogen Concentration Resdiual Plots by Date"}
par(mfrow = c(2,2), mar=c(2,2,2,2))
plot(NitrogenRegressionDate)
par(mfrow = c(1,1))
```

  
  A time series analysis was also completed to see how nitrogen concentration changes over time. The following plot displays the decomposed nitrogen concentration components of the time series run. As you can see, there is seasonality, which confirms our rejecting of the null hypothesis. 
```{r, message=FALSE, fig.cap="Nitrogen Concentration Residual Plots"}
#6. Season Plot #average the months
Nitrogen.monthly.preliminary <- NitrogenYears %>%
  mutate(CleanDate = my(paste0(Month,"-",Year)))

Nitrogen.monthly <-aggregate(Nitrogen.monthly.preliminary$Nitrogen_Percent,  
                                  by=list(Nitrogen.monthly.preliminary$CleanDate),  
                                  FUN=mean)

colnames(Nitrogen.monthly) <- c("Date", "Mean_NitrogenPercent")

#7. Generate time series monthly
f_monthmonthly <- month(first(Nitrogen.monthly$Date))
f_yearmonthly <- year(first(Nitrogen.monthly$Date))

Nitrogen.monthly.ts <- ts(Nitrogen.monthly$Mean_NitrogenPercent,
                   start=c(f_yearmonthly,f_monthmonthly),
                   frequency=12) 

#Decompose monthly
Nitrogen.monthly.ts.decomp <- stl(Nitrogen.monthly.ts,s.window = "periodic")
plot(Nitrogen.monthly.ts.decomp)


```
  
  The Mann-Kendall analysis performed on the nitrogen concentration non-seasonal time series produced a p-value of less than 0.05 (<.001, tau value of -0.19). As a result, the data displays a significant, decreasing trend for nitrogen concentration. The plot below displays the non-seasonal nitrogen concentration amount in the time series which was produced by removing the seasonal component from the time series.
```{r, message=FALSE, fig.cap="Time Series Analysis for Nitrgoen Concentration"}
#12 Run SMK test monthly
Nitrogen_Monthly_SMK <- Kendall::SeasonalMannKendall(Nitrogen.monthly.ts)

Nitrogen_Monthly_SMK
summary(Nitrogen_Monthly_SMK)

# 13 Visualization
Nitrogen_Plot <-
ggplot(Nitrogen.monthly, aes(x = Date, y = Mean_NitrogenPercent)) +
  geom_point() +
  geom_line() +
  labs("Mean Nitrogen (percent)", title="Nitrogen Concentration Percent over Time") +
  geom_smooth( method = lm )

print(Nitrogen_Plot)

ggsave(file=here('Output', 'Analysis','NitrogenConcentrationPointbySeason.png'), width = 10, height = 10, units = "cm")


#taking out seasonal
NitrogenComponents <- as.data.frame(Nitrogen.monthly.ts.decomp$time.series[,1:3])

Nitrogen_NoSeasonal <-NitrogenComponents$trend+NitrogenComponents$remainder

#16 Run MK test monthly
Nitrogen_Monthly_MK <- Kendall::MannKendall(Nitrogen_NoSeasonal)

Nitrogen_Monthly_MK
summary(Nitrogen_Monthly_MK)
```
\newpage 

## Question 2: Does giant kelp nitrogen concentration amount vary by season?  

For the second question, we wanted to know the relationship between nitrogen concentration and season. Winter is defined as beggining of December through February, Spring is beginning of March through May, Summer is June through August, and Fall is September through November. 

H0 : There is no significant difference of nitrogen concentration by season.
HA : There is significant difference of nitrogen concentration by season.

The interaction between nitrogen concentration and season is significant because the p-value is < .05  (p < 2.2e-16). We can reject the null hypothesis and accept that there is significant difference of nitrogen concentration by season.The box plot below shows the change in nitrogen concentration over seasons.
```{r, echo=FALSE, message=FALSE}
NitrogenYears$Month <- factor(NitrogenYears$Month, levels = c(1,2,3,4,5,6,7,8,9,10,11,12),labels =c(  
    "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep",  
    "Oct", "Nov", "Dec") )

NitrogenYears <- NitrogenYears %>%
  mutate(SeasonNames = fct_collapse(Month,
    "Spring" = month.abb[3:5],
    "Summer" = month.abb[6:8],
    "Fall" = month.abb[9:11],
    "Winter" = month.abb[c(12,1,2)]))


NitrogenRegressionSeason <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$SeasonNames)
#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .75 on 1280 degrees of freedom. However, year only explains 5% of nitrogen percent amount. 
summary(NitrogenRegressionSeason)

#NitrogenRegressionMonth <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Month)
#summary(NitrogenRegressionMonth)


```
  
    Below are the normal distribution residuals charts for the linear regression of nitrogen concentration by season.

```{r, message=FALSE, fig.cap="Nitrogen Concentration Residuals Plots for Season"}

#3. Normal Distribution
par(mfrow = c(2,2), mar=c(2,2,2,2))
plot(NitrogenRegressionSeason)
par(mfrow = c(1,1))

```

```{r, message=FALSE, fig.cap="Nitrgoen Concentration over Seasons"}

NitrogenSeasonPlot <- ggplot(NitrogenYears, aes(x=SeasonNames, y=Nitrogen_Percent))+
  geom_boxplot()+
  labs(title="Nitrogen Concentration Percent by Season", x="Season", y="Nitrogen Concentration Percent")

print(NitrogenSeasonPlot)

ggsave(file=here('Output', 'Analysis','NitrogenConcentrationBoxbySeason.png'), width = 10, height = 10, units = "cm")

```

\newpage
## Question 3:  Does nitrogen surge uptake vary between the spring and summer seasons?
The last three questions focus on the experiment completed over five days in the spring and summer of 2019.  Nitrogen was blasted for 45 minutes and data on nitrogen and biomass update was recorded. 



H0 : There is no significant difference of nitrogen surge uptake between spring and summer seasons.
HA : There is significant difference of nitrogen surge uptake between spring and summer seasons.

The interaction between nitrogen uptake and season is significant because the p-value is < .05 (p < .0216). We can reject the null hypothesis and accept that there is significant difference of nitrogen surge uptake between spring and summer seasons.. The R-squared is .014, so a low amount of nitrogen uptake variance is in response to data being collected in the spring or summer. The box plot below shows the change in nitrogen concentration over seasons.

```{r, message=FALSE, fig.cap="Nitrogen Uptake Rate by Season Box Plot"}
NitrogenSeasonBox <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Season, y = Uptake_Rate)) +
  geom_boxplot()+
  labs(title="Nitrogen Uptake Rate by Season", x="Season", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenSeasonBox)
#ggsave(file=here('Output', 'Analysis','NitrogenBoxSeason.png'), width = 10, height = 10, units = #"cm")

```
  
  The scatter plot below shows the change in nitrogen concentration over seasons by testing day replicate.
```{r, message=FALSE, fig.cap="Nitrogen Uptake by Season Scatter Plot"}

NitrogenSeasonPoint <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Day_Replicate, y = Uptake_Rate, shape=Season)) +
  geom_point(alpha=.3)+
  labs(title="Nitrogen Uptake Rate by Day and Season", x="Days", y="Nitrogen Uptake Rate (V(hr1))")


print(NitrogenSeasonPoint)
#ggsave(file=here('Output', 'Analysis','NitrogenPointSeason.png'), width = 10, height = 10, units = #"cm")
```

```{r, message=FALSE}

#linear regression
NitrogenRegressionSeason <- lm(SurgeUptakeNitrogen$Uptake_Rate ~ SurgeUptakeNitrogen$Season)

summary(NitrogenRegressionSeason)
```

\newpage
## Question 4: Does giant kelp nitrogen surge uptake vary based on the amount of exposure?
H0 : There is no significant difference of nitrogen surge uptake between 
different exposure lengths.
HA : There is significant difference of nitrogen surge uptake between different
exposure lengths.

The interaction between nitrogen uptake and season is significant because the p-value is < .05 (p = 8.394e-06). We can reject the null hypothesis and accept that there is significant difference of nitrogen surge uptake between different exposure lengths. The R-squared is .014, so a low amount of nitrogen uptake variance is in response to data being collected in the spring or summer. The box plot below shows the change in nitrogen concentration over seasons.

The Tukey HSD test shows that there is significant difference between the means of time interval 1 and 45 (p=value = 0.0000068), intervals 1 and 15 (p-value=.00076), and intervals 1 and 5 (p-value = .04). There is not significant difference between the means of time interval 15 and 45 (p-value = .71), 5 and 45 (p-value = .10), and 5 and 15 (p-value = .62) because their p-value is >.05. 

```{r, echo=FALSE, message=FALSE, fig.cap="Nitrogen Uptake Rate by Exposure Length Box Plot"}
#can only put one plot at a time because of figure titles

ExposureLength <- SurgeUptakeNitrogen %>%
  mutate(Day_Replicate = as.factor(Day_Replicate))%>%
  mutate(Time = as.factor(Time))


ExposureLength$Season <-factor(SurgeUptakeNitrogen$Season, labels = c("Spring", "Summer"))
ExposureLength$Nitrogen_Types<-factor(SurgeUptakeNitrogen$Nitrogen_Types, labels= c("Ammonium", "Nitrate", "Urea"))


NitrogenTimeBox <- 
  ggplot(ExposureLength, aes(x = Time, y = Uptake_Rate)) +
  geom_boxplot()+
  labs(title="Nitrogen Uptake Rate by Exposure Length", x="Exposure Length (minutes)", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenTimeBox)
ggsave(file=here('Output', 'Analysis','NitrogenTimeBox.png'), width = 10, height = 10, units = "cm")


```
```{r, message=FALSE, echo=FALSE, fig.cap="Nitrogen Uptake Rate by Day Replicate Scatter Plot"}
NitrogenTimePointbyDay <- 
  ggplot(ExposureLength, aes(x = Day_Replicate, y = Uptake_Rate, shape=Time)) +
  geom_point(alpha=.2)+
  labs(title="Nitrogen Uptake Rate by Day and Exposure Length", x="Days", y="Nitrogen Uptake Rate (V(hr1))")


print(NitrogenTimePointbyDay)
ggsave(file=here('Output', 'Analysis','NitrogenTimePointbyDay.png'), width = 10, height = 10, units = "cm")
```


```{r, echo=FALSE, message=FALSE, fig.cap="Nitrogen Uptake Rates by Type with Time Scatter Plot"}
NitrogenTimePointbyType <- 
  ggplot(ExposureLength, aes(x = Nitrogen_Types, y = Uptake_Rate, shape=Time)) +
 geom_point(alpha=.2)+
 labs(title="Nitrogen Uptake Rate by Types and Exposure Length", x="Nitrogen Types", y="Nitrogen Uptake Rate (V(hr1))")

print(NitrogenTimePointbyType)
ggsave(file=here('Output', 'Analysis','NitrogenTimePointbyType.png'), width = 10, height = 10, units = "cm")
```

```{r, message=FALSE}
TimeLM <-lm(data = ExposureLength, Uptake_Rate~Time)
summary(TimeLM)

TimeAOV <- aov(data=ExposureLength, Uptake_Rate~Time)
summary(TimeAOV)


#TimeNitrogenAIC <- lm(data=ExposureLength, Uptake_Rate~Time+Nitrogen_Types+Season)
#step(TimeNitrogenAIC)


TimeDurationAOV <- aov(data=ExposureLength, Uptake_Rate~Time)
summary(TimeDurationAOV)

TukeyHSD(TimeDurationAOV)
```


\newpage
## Question 5: Does uptake vary by type of nitrogen (ammonium, nitrate, and urea)?
H0 : There is no significant difference of nitrogen surge uptake of different nitrogen types.
HA : There is significant difference of nitrogen surge uptake between different nitrogen types.

The interaction between nitrogen uptake and season is significant because the p-value is < .05 (p < 2.2e-16). We can reject the null hypothesis and accept that there is significant difference of nitrogen surge uptake between spring and summer seasons.. The R-squared is .192, so a moderate amount of nitrogen uptake variance is in response to the nitrogen type being blasted. The box plot below shows the change in nitrogen uptake rate by nitrogen type.

The Tukey HSD test shows that there is significant difference between the means of time interval 1 and 45 (p=value = 0.0000068), intervals 1 and 15 (p-value=.00076), and intervals 1 and 5 (p-value = .04). There is not significant difference between the means of time interval 15 and 45 (p-value = .71), 5 and 45 (p-value = .10), and 5 and 15 (p-value = .62) because their p-value is >.05.

```{r, message=FALSE, fig.cap="Nitrogen Uptake by Type Box Plot"}
# Creating a ggplot to compare uptake of various types of nitrogen

NitrogenUptaketype <- 
  ggplot(SurgeUptakeNitrogen, aes(x = Nitrogen_Types, y = Uptake_Rate)) +
  geom_boxplot() +
  labs(title="Uptake Rate by Nitrogen Type", x="Nitrogen Types", y="Nitrogen Uptake Rate (V(hr1))")
print(NitrogenUptaketype)
ggsave(file=here('Output', 'Analysis','NitrogenUptakeBoxtype.png'), width = 10, height = 10, units = "cm")
# Creating a ggplot to compare biomass uptake of various types of nitrogen

```


The jitter plot below shows the change in nitrogen uptake rate by nitrogen type with day replicate added.

```{r, message=FALSE, fig.cap="Nitrogen Uptake by Type Jigger Plot"}

# Creating a ggplot to compare how nitrogen uptake varies per day per nitrogen type
NitrogenDayUptake <-
  ggplot(SurgeUptakeNitrogen, aes(x = Nitrogen_Types, y = Uptake_Rate, shape = Day_Replicate)) +
  geom_jitter(alpha = 0.5) +
#  scale_color_manual(values = c("#FF0000", "#BF003F", "#7F007F", "#3F00BF", #"#0000FF"))+
  labs(title="Nitrogen Uptake per Day for Each Nitrogen Type", 
       x="Nitrogen Types", 
       y="Nitrogen Uptake Rate (V(hr1))", color = "Day")

print(NitrogenDayUptake)
ggsave(file=here('Output', 'Analysis','NitrogenUptakeJitterType.png'), width = 10, height = 10, units = "cm")

```



```{r, message=FALSE}
# Generalized Linear Model (GLM)
NitorgenLM <-lm(data = SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types)
summary(NitorgenLM)
```


```{r}
# One-way ANOVA
NitrogenAOV <- aov(data = SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types)
summary(NitrogenAOV)
TukeyHSD(NitrogenAOV)
```


```{r}
# One-way ANOVA
NitrogenAOV <- aov(data = SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types)
summary(NitrogenAOV)
TukeyHSD(NitrogenAOV)

# AIC-based Model Selection:
NitrogenTypeAIC <- lm(data=SurgeUptakeNitrogen, Uptake_Rate~Nitrogen_Types+Season+Time)
step(NitrogenTypeAIC)
summary(NitrogenTypeAIC)

# For biomass
#BiomassTypeAIC <- lm(data=SurgeUptakeNitrogen, Biomass_Uptake~Nitrogen_Types+Season+Time)
#step(NitrogenTypeAIC)
#summary(NitrogenTypeAIC)
```

\newpage

# Summary and Conclusions

Akaike information criterion (AIC) was performed for the final three questions to see if those three variables all contributed to the variation of nitrogen uptake. All three variables (season, nitrogen exposure lengths, and nitrogen type), contributed to the variance of nitrogen uptake rate. 

What does these p-values about the first data set tell us about the intro

What does the second set  tell us about the intro as well as the data set (aka, more testing needs to be done, was a short experiment)

Table: Significance Values

Research Question | P-Value
-----:|:----:
Question 1| < 2.2e-16
Question 2| < 1.16e-4
Question 3| .0216
Question 4| 
Question 5| < 2.2e-16
\newpage

# References

Cedeno, T., Brzezinski, M., Miller, R., Reed, D. (2021). Santa Barbara Coastal LTER: Reef: Surge uptake capability in Macrocystis pyrifera in response to pulses of three different forms of nitrogen ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/d079ff0e95defb71b87040f56235d755 

Conroy, G. (2023).  Why Earth’s giant kelp forests are worth $500 billion a year. Nature. https://www.nature.com/articles/d41586-023-01307-3 

Fernández, P.A., Gaitán-Espitia, J.D., Leal, P.P. et al. (2020). Nitrogen sufficiency enhances thermal tolerance in habitat-forming kelp: implications for acclimation under thermal stress. Scientific Reports 10, 3186.https://doi.org/10.1038/s41598-020-60104-4

Hurlimann, S. (2019, July 4). How kelp naturally combats global climate change. Science in the News.https://sitn.hms.harvard.edu/flash/2019/how-kelp-naturally-combats-global-climate-change/  


Legendre, L., Gosselin, M., (1997). Estimation of N or C uptake rates by phytoplankton using 15N or 13C: revisiting the usual computation formulae. Journal of Plankton Research 19, Issue 2, Pages 263–271. https://academic.oup.com/plankt/article/19/2/263/1544362. 


Wilson, S. (2019). Fires, floods and free parking: California’s unending fight against climate change. The Washington Post. https://www.washingtonpost.com/graphics/2019/national/climate-environment/climate-change-california/

Santa Barbara Coastal LTER, Reed, D., Miller, R. (2021). SBC LTER: Reef: Macrocystis pyrifera CHN content (carbon, hydrogen, nitrogen), ongoing since 2002 ver 21. Environmental Data Initiative. https://doi.org/10.6073/pasta/48686fef076a23310a0e48f69d76260c (Accessed 2023-12-05).

