---
title: "Trends in giant kelp nitrogen surge uptake"
author: "Erin Ansbro, Cara Kuuskvere, Tani Valdez Rivas"
date: "December 2023"
output: 
  html_document:
    toc: true
---
\newpage
\tableofcontents
\newpage
\listoftables 
\newpage
\listoffigures 
\newpage

```{r include=FALSE}
# Set your working directory
library(here)
here()

# Loading packages
library(tidyverse)
library(lubridate)
library(cowplot)
library(knitr)
library(sf)
library(leaflet)
library(mapview); mapviewOptions(fgb = FALSE)

# Reading in the data
SurgeUptakeNitrogen <-
  read.csv(here('Data','Processed','Nitrogen_Processed.csv'), stringsAsFactors=TRUE)

# Set your ggplot theme
#Create and set theme 
mytheme <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
theme_set(mytheme)

# Load your datasets
SurgeUptakeNitrogen <-
  read.csv(here('Data','Processed','Nitrogen_Processed.csv'), stringsAsFactors=TRUE)

NitrogenYears <-read.csv(here('Data','Processed','NitrogenYears_Processed.csv'), stringsAsFactors=TRUE)

#Verifying data attributes
SurgeUptakeNitrogen$Time <- as.factor(SurgeUptakeNitrogen$Time)
SurgeUptakeNitrogen$Blade_Replicate <- as.factor(SurgeUptakeNitrogen$Blade_Replicate)
SurgeUptakeNitrogen$Day_Replicate <- as.factor(SurgeUptakeNitrogen$Day_Replicate)

```


# Rationale and Research Questions

Kelp forests along the United States West Coast are impacted by increasing temperatures as a result of climate change. Since the late 19th century, Santa Barbara County in Southern California has had an increase in average temperature by 4.1 degrees Fahrenheit (Wilson, 2019). Kelp not only provides a healthy ecosystem for other flora and fauna but also provides ecosystem services to humans as well. Kelp assists with human-related environmental issues in two ways. First, kelp can sequester a large amount of carbon, aiding in climate change mitigation (Hurlimann, 2019). Secondly, they absorb excess nitrogen, which can lessen the impacts of agricultural runoff (Conroy, 2023). 
A recent study has shown that increased nitrogen uptake by kelp can limit the effect that heat has on them. This is an interesting development because nitrogen and eutrophication are generally seen as negative consequences of human interference. However, in this case, it could assist kelp and help them survive increased climate-change-related temperature increases (Fernández, 2020).
Interested in this recent research, data about nitrogen concentration amount and uptake were located in the Long-Term Ecological Research database titled “SBC LTER: Reef: Surge uptake capability in Macrocystis pyrifera in response to pulses of three different forms of nitrogen” (Cedeno, 2021) and ” SBC LTER: Reef: Macrocystis pyrifera CHN content (carbon, hydrogen, nitrogen), ongoing since 2002” (Santa Barbara, 2021). This data looked simple enough to clean and manipulate, while also providing enough information for a research project. 

We are interested in trends in giant kelp nitrogen concentration amount and nitrogen surge uptake. Specific questions include:
1.	Has nitrogen concentration in kelp changed over the years (2002-2021)?
2.	Does giant kelp nitrogen concentration amount vary by season?
3.	Does nitrogen surge uptake vary between the spring and summer seasons?
4.	Does giant kelp nitrogen surge uptake vary based on the amount of exposure?
5.	Does uptake vary by type of nitrogen (ammonium, nitrate, and urea)?

Github Information: https://github.com/Tani-ValdezRivas/ValdezKuuskvereAnsbro


\newpage

# Dataset Information

Data sets were found on the Environmental Data Initiative. All sets were from the Santa Barbara Coastal Long Term Ecological Research Site (SBC LTER) which is based  within the University of California, Santa Barbara (UCSB) Marine Science Institute, and is part of the National Science Foundation's LTER Network. 

The first data set focuses on one experiment done that examines nitrogen and biomass uptake by kelp during timed nitrogen blasting.  The first data set was used to determine if nitrogen uptake varies by time (minutes), season, or nitrogen type. The second data set has dates and percent nitrogen of kelp over almost a twenty year period. This second data set was used to gain a general sense of nitrogen found in kelp in Santa Barbara, CA and to see if that has changed over time.

Table: Data Set Information

Detail | Data Set 1 | Data Set 2
-----:|:----:|:----:
Source| Santa Barbara Coastal Long Term Ecological Research (LTER) | Santa Barbara Coastal  Long Term Ecological Research (LTER)
Retrieved from | https://portal.edirepository.org/nis/mapbrowse?packageid=edi.1022.2 |https://sbclter.msi.ucsb.edu/data/catalog/package/?package=knb-lter-sbc.24
Variables Used | Nitrogen Types, Season, Day Replicate, Time, Blade Replicate, Uptake Rate, Biomass Uptake Rate |Year, Month, Date, Nitrogen Percent
Date | 2021 | 2002-2021

The data for nitrogen uptake was cleaned by renaming the column names to be better understood by the public. Multiple columns including columns for Blade Replicate, Day Replicate, Time, Season, and Nitrogen Type were changed to factors. Season and Nitrogen Type also had labels added to capitalize the names of seasons and nitrogen types.Data cleaning for the second data set entailed renaming column names as well. Date was changed from a factor to a date data type. Sites with missing values were also removed. 

\newpage

# Exploratory Analysis 

A map was created to gain a sense of location. The Map shows the sampling sites of nitrogen uptake along the coast of California.
```{r Map, echo=FALSE, message=FALSE}

# Data Exploration Part 1: Mapping the locations of the study sites

# Creating a data frame of the locations
Site_Name <- c("Mohawk Reef", "Arroyo Quemado Reef", "Arroyo Burro Reef")
Longitude <- c("-119.7295", "-120.11905", "-119.7445915	")
Latitude <- c("34.3940708", "34.46774988", "34.400275")

Site_information <- data.frame(Site_Name, Longitude, Latitude)

# Changing Site Name, Longitude, Latitude to factors and numerics
Site_information$Site_Name <- as.factor(Site_information$Site_Name)
Site_information$Longitude <- as.numeric(Site_information$Longitude)
Site_information$Latitude <- as.numeric(Site_information$Latitude)


# Converting the dataframe into a spatial reference frame
Surge_Reef_sf <- Site_information %>% 
  st_as_sf(coords = c('Longitude','Latitude'),
           crs=4326)

# Making a map of the study sites
Surge_Reef_Map <- mapview(Surge_Reef_sf, layer.name = 'Reef Site locations')
Surge_Reef_Map
```
Boxplot text
```{r Box Plots, echo=FALSE, message=FALSE}
# Data Exploration Part 2.1: Box Plots to visualize data means and quartiles
UptakeRateBoxbyType <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Nitrogen_Types, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Nitrogen Type")
UptakeRateBoxbyType
ggsave(file=here('Output', 'BoxDataExp','UptakeRateBoxbyType.png'), width = 10, height = 10, units = "cm")

BiomassBoxbyType <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Nitrogen_Types, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Nitrogen Type")
BiomassBoxbyType
ggsave(file=here('Output','BoxDataExp', 'BiomassBoxbyType.png'), width = 10, height = 10, units = "cm")

UptakeRateBoxbySeason <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Season, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Season")
UptakeRateBoxbySeason
ggsave(file=here('Output', 'BoxDataExp','UptakeRateBoxbySeason.png'), width = 10, height = 10, units = "cm")

BiomassBoxbySeason <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Season, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Season")
BiomassBoxbySeason
ggsave(file=here('Output','BoxDataExp', 'BiomassBoxbySeason.png'), width = 10, height = 10, units = "cm")

UptakeRateBoxbyTime <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Time, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Length of Exposure")
UptakeRateBoxbyTime
ggsave(file=here('Output','BoxDataExp', 'UptakeRateBoxbyTime.png'), width = 10, height = 10, units = "cm")

BiomassBoxbyTime <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Time, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Length of Exposure")
BiomassBoxbyTime
ggsave(file=here('Output','BoxDataExp', 'BiomassBoxbyTime.png'), width = 10, height = 10, units = "cm")

UptakeRateBoxbyDayRep <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Day_Replicate, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Day Replicate")
UptakeRateBoxbyDayRep
ggsave(file=here('Output', 'BoxDataExp','UptakeRateBoxbyDayRep.png'), width = 10, height = 10, units = "cm")

BiomassBoxbyDayRep <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Day_Replicate, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Day Replicate")
BiomassBoxbyDayRep
ggsave(file=here('Output','BoxDataExp', 'BiomassBoxbyDayRep.png'), width = 10, height = 10, units = "cm")

UptakeRateBoxbyBladeRep <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Blade_Replicate, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Blade Replicate")
UptakeRateBoxbyBladeRep
ggsave(file=here('Output','BoxDataExp', 'UptakeRateBoxbyBladeRep.png'), width = 10, height = 10, units = "cm")

BiomassBoxbyBladeRep <- ggplot(SurgeUptakeNitrogen) +
  geom_boxplot(aes(x = Blade_Replicate, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Blade Replicate")
BiomassBoxbyBladeRep
ggsave(file=here('Output', 'BoxDataExp','BiomassBoxbyBladeRep.png'), width = 10, height = 10, units = "cm")
```


Violin text
```{r Violins, echo=FALSE, message=FALSE}
# Data Exploration Part 2.2: Violins to visualize range and spread of each component of the data set
UptakeRateViolinbyType <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Nitrogen_Types, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Nitrogen Type")
UptakeRateViolinbyType
ggsave(file=here('Output','ViolinDataExp', 'UptakeRateViolinbyType.png'), width = 10, height = 10, units = "cm")

BiomassViolinbyType <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Nitrogen_Types, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Nitrogen Type")
BiomassViolinbyType
ggsave(file=here('Output', 'ViolinDataExp','BiomassViolinbyType.png'), width = 10, height = 10, units = "cm")

UptakeRateViolinbySeason <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Season, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Season")
UptakeRateViolinbySeason
ggsave(file=here('Output','ViolinDataExp', 'UptakeRateViolinbySeason.png'), width = 10, height = 10, units = "cm")

BiomassViolinbySeason <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Season, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Season")
BiomassViolinbySeason
ggsave(file=here('Output','ViolinDataExp', 'BiomassViolinbySeason.png'), width = 10, height = 10, units = "cm")

UptakeRateViolinbyTime <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Time, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Length of Exposure")
UptakeRateViolinbyTime
ggsave(file=here('Output', 'ViolinDataExp','UptakeRateViolinbyTime.png'), width = 10, height = 10, units = "cm")

BiomassViolinbyTime <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Time, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Length of Exposure")
BiomassViolinbyTime
ggsave(file=here('Output', 'ViolinDataExp','BiomassViolinbyTime.png'), width = 10, height = 10, units = "cm")

UptakeRateViolinbyDayRep <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Day_Replicate, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Day Replicate")
UptakeRateViolinbyDayRep
ggsave(file=here('Output', 'ViolinDataExp','UptakeRateViolinbyDayRep.png'), width = 10, height = 10, units = "cm")

BiomassViolinbyDayRep <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Day_Replicate, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Day Replicate")
BiomassViolinbyDayRep
ggsave(file=here('Output', 'ViolinDataExp','BiomassViolinbyDayRep.png'), width = 10, height = 10, units = "cm")

UptakeRateViolinbyBladeRep <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Blade_Replicate, y = Uptake_Rate))+
  labs(title="Uptake Rate (V(hr1)) by Blade Replicate")
UptakeRateViolinbyBladeRep
ggsave(file=here('Output', 'ViolinDataExp','UptakeRateViolinbyBladeRep.png'), width = 10, height = 10, units = "cm")

BiomassViolinbyBladeRep <- ggplot(SurgeUptakeNitrogen) +
  geom_violin(aes(x = Blade_Replicate, y = Biomass_Uptake))+
  labs(title= "Biomass Uptake by Blade Replicate")
BiomassViolinbyBladeRep
ggsave(file=here('Output', 'ViolinDataExp','BiomassViolinbyBladeRep.png'), width = 10, height = 10, units = "cm")

```
The last part of the exploratory analysis focused on an overview of nitrogen concentration amount over the last twenty years. A summary was used on the entire data set as well as on the nitrogen concentration column to gain a general sense of what amounts were present.  
```{r}
#Data Exploration Part 3

summary(NitrogenYears)

summary(NitrogenYears$Nitrogen_Percent)
```
A scatter plot (Fig 1.2) was used to visualize the relationship between year and nitrogen concentration amount.
```{r}
#Data Exploration Part 3

NitrogenYears_Point <- ggplot(NitrogenYears, aes(x=Year, y=Nitrogen_Percent))+
  geom_point(alpha=.5)+
  labs(title="Nitrogen Percent by Year", x="Date", y="Nitrogen Percent")

print(NitrogenYears_Point)
```

Lastly, a boxplot, see below (Fig 1.3) was used to visualize the relationship between sites and nitrogen concentration. Looking at this data, it appeared concentration amount was not too different by site. Looking at the scatter plot for year, it looked like nitrogen decreased slightly. This was then explored further in the Data Analysis section.

```{r Nitrogen Data Exploration, echo=FALSE, message=FALSE}
#Data Exploration Part 3

NitrogenYears_Box <- ggplot(NitrogenYears, aes(x=NewSites, y=Nitrogen_Percent))+
  geom_boxplot()+
  labs(title="Nitrogen Percent by Site", x="Site", y="Nitrogen Percent")

print(NitrogenYears_Box)
```

\newpage

# Analysis


## Question 1: Does giant kelp nitrogen percent vary over year or season?

For the first question, we wanted to understand the relationship between nitrogen concentration and time. Linear regression and normal distribution tests were run for both the year and the specific date of collection. The boxplot (Fig 2.1) below shows the nitrogen concentrations over the years, starting with 2002. 

H0 : There is no significant difference of nitrogen percent over the past twenty years.
HA : There is significant difference of nitrogen percent over the past twenty years.

The interaction between nitrogen concentration and year is significant (p < 0.001). However, from the ___the year only explains 5% of the nitrogen percent amount. A correlation test was also run, with the score being -.202. There is a negative, somewhat weak correlation between nitrogen percent and year. 
Due to this, we then moved to date instead of year to see if that increased the ___. The interaction between nitrogen concentration and collection date is significant (p < 0.001). We can reject the null hypothesis and accept that nitrogen concentration changes _____. The date explains 97.4% of nitrogen percent amount change. Still, we can reject the null hypothesis and accept that the date predicts nitrogen concentration. Below is a boxplot showing change in nitrogen concentration over time (Fig).


```{r}
#1. LM for year
NitrogenRegressionYear <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Date)
#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .75 on 1280 degrees of freedom. However, year only explains 5% of nitrogen percent amount. 
summary(NitrogenRegressionYear)
#Correlation
cor.test(NitrogenYears$Nitrogen_Percent, NitrogenYears$Year)
#Running correlation test, the correlation score is -.202. So there is a negative correlation. So a somewhat weak correlation

#2. LM for Date
NitrogenRegressionDate <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Date)

#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .1648 on 728 degrees of freedom. Date explains 97.4% of nitrogen percent amount. 
summary(NitrogenRegressionDate)

#can't run corr because it is a date
#cor.test(NitrogenYears$Nitrogen_Percent, NitrogenYears$Date)


NitrogenYears$Year <- as.factor(NitrogenYears$Year)
NitrogenPlot <- ggplot(NitrogenYears, aes(x=Year, y=Nitrogen_Percent))+
  geom_boxplot()
print(NitrogenPlot)
```

Below are the normal distribution residuals charts for the linear regression of change nitrogen by date.(Fig 2.2)
```{r}
par(mfrow = c(2,2), mar=c(2,2,2,2))
plot(NitrogenRegressionDate)
par(mfrow = c(1,1))
```

A time series analysis was completed to see how nitrogen concentration changes over time. The following plot displays the decomposed nitrogen concentration components of the time series run (Fig 2.3). As you can see, there is seasonality, which confirms our rejecting of the null hypothesis. 
```{r}
#6. Season Plot #average the months
Nitrogen.monthly.preliminary <- NitrogenYears %>%
  mutate(CleanDate = my(paste0(Month,"-",Year)))

Nitrogen.monthly <-aggregate(Nitrogen.monthly.preliminary$Nitrogen_Percent,  
                                  by=list(Nitrogen.monthly.preliminary$CleanDate),  
                                  FUN=mean)

colnames(Nitrogen.monthly) <- c("Date", "Mean_NitrogenPercent")

#7. Generate time series monthly
f_monthmonthly <- month(first(Nitrogen.monthly$Date))
f_yearmonthly <- year(first(Nitrogen.monthly$Date))

Nitrogen.monthly.ts <- ts(Nitrogen.monthly$Mean_NitrogenPercent,
                   start=c(f_yearmonthly,f_monthmonthly),
                   frequency=12) 

#Decompose monthly
Nitrogen.monthly.ts.decomp <- stl(Nitrogen.monthly.ts,s.window = "periodic")
plot(Nitrogen.monthly.ts.decomp)


```
The Mann-Kendall analysis performed on the nitrogen concentration non-seasonal time series produced a tau value of -0.19 with a p-value of less than 0.05. As a result, the data displays a significant, decreasing trend for nitrogen concentration. The plot below displays the non-seasonal nitrogen concentration amount in the time series which was produced by removing the seasonal component from the time series (Fig 2.4).
```{r}
#12 Run SMK test monthly
Nitrogen_Monthly_SMK <- Kendall::SeasonalMannKendall(Nitrogen.monthly.ts)

Nitrogen_Monthly_SMK
summary(Nitrogen_Monthly_SMK)

# 13 Visualization
Nitrogen_Plot <-
ggplot(Nitrogen.monthly, aes(x = Date, y = Mean_NitrogenPercent)) +
  geom_point() +
  geom_line() +
  ylab("Mean Nitrogen (percent)") +
  geom_smooth( method = lm )

print(Nitrogen_Plot)

#taking out seasonal
NitrogenComponents <- as.data.frame(Nitrogen.monthly.ts.decomp$time.series[,1:3])

Nitrogen_NoSeasonal <-NitrogenComponents$trend+NitrogenComponents$remainder

#16 Run MK test monthly
Nitrogen_Monthly_MK <- Kendall::MannKendall(Nitrogen_NoSeasonal)

Nitrogen_Monthly_MK
summary(Nitrogen_Monthly_MK)
```


## Question 2: Does giant kelp nitrogen concentration amount vary by season?
For the second question, we wanted to know the relationship between concentration and season. 

H0 : There is no significant difference of nitrogen concentration by season.
HA : There is significant difference of nitrogen concentration by season.

The interaction between nitrogen concentration and collection date is significant (p < 0.001). We can reject the null hypothesis and accept that nitrogen concentration changes _____.
The boxplot below shows the change in nitrogen concentration over seasons (Fig 2.5). 

```{r}
NitrogenYears$Month <- factor(NitrogenYears$Month, levels = c(1,2,3,4,5,6,7,8,9,10,11,12),labels =c(  
    "Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep",  
    "Oct", "Nov", "Dec") )

NitrogenYears <- NitrogenYears %>%
  mutate(SeasonNames = fct_collapse(Month,
    "Spring" = month.abb[3:5],
    "Summer" = month.abb[6:8],
    "Fall" = month.abb[9:11],
    "Winter" = month.abb[c(12,1,2)]))


NitrogenRegressionSeason <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$SeasonNames)
#Linear regression on the nitrogen percent amount by year. P-value is less than .05. Which means there is significant difference. Residual standard error is small at .75 on 1280 degrees of freedom. However, year only explains 5% of nitrogen percent amount. 
summary(NitrogenRegressionSeason)

NitrogenRegressionMonth <- lm(NitrogenYears$Nitrogen_Percent ~ NitrogenYears$Month)
summary(NitrogenRegressionMonth)


```

```{r}

#3. Normal Distribution
par(mfrow = c(2,2), mar=c(2,2,2,2))
plot(NitrogenRegressionSeason)
par(mfrow = c(1,1))

```

```{r}

NitrogenSeasonPlot <- ggplot(NitrogenYears, aes(x=SeasonNames, y=Nitrogen_Percent))+
  geom_boxplot()
print(NitrogenSeasonPlot)
```


## Question 3:  Does nitrogen surge uptake vary between the spring and summer seasons?
The last three questions focus on the experiment completed over five days in the spring and summer of 2019.  Nitrogen was blasted for 45 minutes and data on nitrogen and biomass update was recorded. 


H0 : There is no significant difference of nitrogen surge uptake between spring and summer seasons.
HA : There is significant difference of nitrogen surge uptake between spring and summer seasons.

The metadata indicates that the giant kelp blades studied were collected during the nitrate replete spring and the nitrate-deplete summer. The five replicate experiments were conducted in both spring and summer of 2019, which allows us to find potential differences in the nitrogen surge uptake of the two populations. The experimenters calculate uptake rate using a modified version of an equation by Legendre and Grosselin (1997) V= (nt - n0)/[t(ds-d0)] where nt is the atom percentage of Nitrogen-15 in the blade after incubation; n0 is the average concentration of isotope in the control blades; ds is the atom percentage of Nitrogen-15 in the seawater at the start of the incubation; d0 is the atom percentage of Nitrogen-15 in the nitrate pool before isotope tracer was added; and t is the length of the incubation time in minutes.


```{r}

```


## Question 4: Does giant kelp nitrogen surge uptake vary based on the amount of exposure?
H0 : There is no significant difference of nitrogen surge uptake between 
different exposure lengths.
HA : There is significant difference of nitrogen surge uptake between different
exposure lengths.

```{r}

```

## Question 5: Does uptake vary by type of nitrogen (ammonium, nitrate, and urea)?
H0 : There is no significant difference of nitrogen surge uptake of different nitrogen types.
HA : There is significant difference of nitrogen surge uptake between different nitrogen types.
```{r}

```

\newpage

# Summary and Conclusions


\newpage

# References

Cedeno, T., Brzezinski, M., Miller, R., Reed, D. (2021). Santa Barbara Coastal LTER: Reef: Surge uptake capability in Macrocystis pyrifera in response to pulses of three different forms of nitrogen ver 1. Environmental Data Initiative. https://doi.org/10.6073/pasta/d079ff0e95defb71b87040f56235d755 

Conroy, G. (2023).  Why Earth’s giant kelp forests are worth $500 billion a year. Nature. https://www.nature.com/articles/d41586-023-01307-3 

Fernández, P.A., Gaitán-Espitia, J.D., Leal, P.P. et al. (2020). Nitrogen sufficiency enhances thermal tolerance in habitat-forming kelp: implications for acclimation under thermal stress. Scientific Reports 10, 3186.https://doi.org/10.1038/s41598-020-60104-4

Hurlimann, S. (2019, July 4). How kelp naturally combats global climate change. Science in the News.https://sitn.hms.harvard.edu/flash/2019/how-kelp-naturally-combats-global-climate-change/  


Legendre, L., Gosselin, M., (1997). Estimation of N or C uptake rates by phytoplankton using 15N or 13C: revisiting the usual computation formulae. Journal of Plankton Research 19, Issue 2, Pages 263–271. https://academic.oup.com/plankt/article/19/2/263/1544362. 


Wilson, S. (2019). Fires, floods and free parking: California’s unending fight against climate change. The Washington Post. https://www.washingtonpost.com/graphics/2019/national/climate-environment/climate-change-california/

Santa Barbara Coastal LTER, Reed, D., Miller, R. (2021). SBC LTER: Reef: Macrocystis pyrifera CHN content (carbon, hydrogen, nitrogen), ongoing since 2002 ver 21. Environmental Data Initiative. https://doi.org/10.6073/pasta/48686fef076a23310a0e48f69d76260c (Accessed 2023-12-05).

